Dynamic Walking and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-12-04 paper.tex
Adress reviewer issue concerning the definition of the conifguration space.

\newtheorem{theorem}{Theorem}

@@

Some planning methods for free-climbing robots \cite{bretl2006motion}
can be seen as a general way 
to consider quasi-static multi-step planning. They are not however directly applicable to
humanoid dynamically balanced locomotion.
Other recent contributions to the field of locomotion planning  include algorithms 
considering the dynamics at the planning phase \cite{shkolnik2011bounding}. This leads to 
a growth of algorithmic complexity, particularly costly for high-dimensional
systems such as humanoid robots. To the authors knowledge such techniques have not been
used on humanoid robotic platforms so far.
In our work, we show that it is not always
necessary to take into account dynamics while planning, even though the final
output of our algorithm is a dynamically balanced walk motion. This result was first
presented in \cite{dalibard2011small}.

@@

The main contribution of this work is a  whole-body motion planner for humanoid robots
that computes collision-free walking trajectories, based on
exact models of both robot and environment. It is used to solve manipulation tasks
that may require walking. The first stage of our algorithm uses a
constrained sampling-based motion 
planner and  computes a collision-free  statically balanced path  for a robot which can be fixed or sliding 
on the ground. Another  contribution of this  paper is the formal  proof that
this first path can always be  approximated by a dynamically balanced, collision-free
walking  trajectory. We have implemented this well-grounded 
method, and the results have been validated on the HRP-2 robot.

@@

FIXME: Avoid the mixup due to the use of task and constraints?

This section presents an algorithm for constrained motion planning. 
The problem solved here
differs from classic approaches in two ways:
\begin{enumerate}
\item the set of valid configurations is defined implicitly, as the ones satisfying a given 
  set of inverse kinematics and balance tasks;
\item the goal is also defined implicitly, by additional inverse kinematics tasks.
\end{enumerate}
During global planning, we will consider several types of
tasks and constraints for various reasons:
\begin{itemize}
\item Static balance: the CoM of the robot stays above
  the support polygon center, the two feet have fixed position and
  orientation.
\item End-effector position and orientation: the goals of some problems presented
  in the experimental section of this paper are defined as a specific robot hand pose,
  or a gaze direction.
\item Configuration task: our adaptation of randomized motion planning algorithms 
  uses tasks defined as the distance towards a given configuration in $\mathcal{C}$. 
  This will be detailed in the following section.
\end{itemize}

>>>>

\newtheorem{theorem}{Theorem}
\renewcommand{\labelitemi}{$-$}
\newcommand\goalmanifold{\mathcal{M}_{g}}

@@

Some planning methods for free-climbing robots \cite{bretl2006motion}
can be seen as a general way 
to consider quasi-static multi-step planning. They are not however directly applicable to
humanoid dynamically balanced locomotion.
Other recent contributions to the field of locomotion planning  include algorithms 
considering the dynamics at the planning phase \cite{shkolnik2011bounding}. This leads to 
a growth of algorithmic complexity, particularly costly for high-dimensional
systems such as humanoid robots. To the authors knowledge such techniques have not been
used on humanoid robotic platforms so far.

\textbf{In this work, we show that, under some assumptions, it is
  sufficient to plan in the configuration space in a first step, even
  though the final output of our algorithm is guaranteed to be a
  dynamically balanced walk motion. This result was first presented in
  \cite{dalibard2011small}.}

@@

The main contribution of this work is a whole-body motion planner for
humanoid robots that computes collision-free walking trajectories,
based on exact models of both robot and environment. It is used to
solve manipulation tasks that may require walking. The first stage of
our algorithm uses a sampling-based constrained motion planner and
computes a collision-free statically balanced path for a robot which
can be fixed or sliding on the ground.

Another contribution of this paper is the formal proof that this first
path can always be approximated by a dynamically balanced,
collision-free walking trajectory. We have implemented this
well-grounded method, and the results have been validated on the HRP-2
robot.

@@

\textbf{This section presents an algorithm for constrained motion
  planning in the configuration space $\mathcal{C}$. In the particular
  case of a humanoid robot which is an under-actuated system,
  $\mathcal{C}=\mathcal{Q} \times SE(3)$, where $\mathcal{Q}$
  represents the kinematic tree actuators and $SE(3)$ corresponds to a
  fictitious floating joint attached to the root of the tree. If the
  robot has $n$ actuated DoFs, then $dim(\mathcal{Q})=n$ and
  $dim(\mathcal{C})=n+6$. A configuration $q$ of $\mathcal{C}$ is said
  to be valid iff, besides being collision-free, it lies on a specific
  submanifold $\mathcal{M}$ of $\mathcal{C}$. $\mathcal{M}$ is called
  the planning submanifold.}

The problem solved here differs from classic approaches in two ways:
\begin{enumerate}
\item the set of valid configurations is defined implicitly, as the
  set of collision-free configurations satisfying a given set of
  inverse kinematics balance constraints;
\item the goal submanifold $\goalmanifold$ is also defined implicitly,
  by additional inverse kinematics constraints.
\end{enumerate}
During global planning, we will consider several types of
tasks and constraints for various reasons:
\begin{itemize}
\item Static balance: the CoM of the robot stays above
  the support polygon center, the two feet have fixed position and
  orientation.
\item End-effector position and orientation: the goals of some problems presented
  in the experimental section of this paper are defined as a specific robot hand pose,
  or a gaze direction.
\item Configuration task: our adaptation of randomized motion planning algorithms 
  uses tasks defined as the distance towards a given configuration in $\mathcal{C}$. 
  This will be detailed in the following section.
\end{itemize}
